Towards caring 15-minute neighbourhoods

### 2024-09-18 WGiT/WGiT-pres.Rmd
WGiT presentation

---
title: "Towards caring 15-minute neighbourhoods"
subtitle: "Proximity Planning Congress 2024"
author: "Anastasia Soukhov, Léa Ravensbergen, Lucía Mejía Dorantes, Antonio Páez"
date: "09-10-2024"
output:
  xaringan::moon_reader:
    seal: false
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current% / 10" 
      highlightStyle: github
      highlightLines: true
      ratio: '16:9'
      countIncrementalSlides: false
      beforeInit: "macros.js"
---
```{r xaringan-themer, include=FALSE, warning=FALSE}
#this themer package allows us to change the .css themes through this chunk. Easier than changing the .css directly.
library(xaringanthemer)
style_mono_accent(
  base_color = "#7A003C", #maroon
  link_color = "#5E6A71", #grey
  text_bold_color = "#FDBF57", #gold
  text_slide_number_color = "#FDBF57", #gold
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

@@

- $i$ is a set of parcel point origin locations; $j$ is a set of destination locations.
- $O^_j$ is a number of opportunities at $j$.
- $c_{ij}$ is the travel cost between $i$ and $j$.
- $f(\cdot)$ is an impedance function of $c_{ij}$; within the cumulative opportunity approach, it is a binary function that takes the value of 1 if $c_{ij}$ is less than a selected value

@@

### Typologies of 15-minute caring neighbourhoods

@@

- Lower income households tend to reside in $\color{Green}{\text{very completely caring}}$ or $\color{GreenYellow}{\text{completely caring}}$ neighbourhoods; 
- Those with a `r emo::ji("up_arrow")` proportion of children tend to reside in $\color{orange}{\text{somewhat caring - not complete}}$ neighbourhoods;
- Higher income households in DAs with a `r emo::ji("down_arrow")` of single-parent households tend to reside in $\color{red}{\text{not caring - not complete}}$ neighbourhoods.

@@

Who has resides in $\color{Green}{\text{very completely caring}$ or $\color{GreenYellow}{\text{completely caring}$ neighbourhoods?  (maybe people who cannot afford cars???)
- DAs that have lower median household incomes and a smaller children population (lower than 24% of the population are children 14 and under).
- and DAs which have especially high LICO are often in $\color{Green}{\text{very completely caring}$ neighbourhoods.

Who more often resides in $\color{BurntOrange}{\text{somewhat caring - not complete}$ neighbourhoods?  
- Of bottom income DAs, DAs with a high proportion of children population (higher than 24% of the population are children 14 and under).

Who more often resides in $\color{BrickRed}{\text{not caring - not complete}$ neighbourhoods?  
- DAs of higher income DAs and with lower proportion of single-parent households (13% or lower)

@@

### Conclusions: classifying and profiling caring 15-minute neighbourhoods

@@

#### What we did:

- 15 minute-city is bridged with "Mobility of Care": 15-minute caring neighbourhoods.
- A data-driven pattern recognition method (Self-organizing maps) to create **15-minute caring neighbourhood** typologies using **accessibility** and **diversity** indicators is presented.
- Using Self-Organizing Maps again; each parcel tagged with a neighbourhood typology was profiled based on all relevant census variables.
]

@@

#### What we found and future considerations:

- **Gentrification concerns**: populations residing in completely caring neighbourhoods are often lower-income and vice-versa.

- **Concerning caring for children and children's self-care**: Lower-income DAs with a `r emo::ji("up_arrow")` proportion of children, and DAs with `r emo::ji("up_arrow")` proportion single-parent households have medium complete care access, why?

@@

>>>>

---
title: "Towards caring 15-minute neighbourhoods"
subtitle: "Proximity Planning Congress 2024"
author: "Anastasia Soukhov, Léa Ravensbergen, Lucía Mejía Dorantes, Antonio Páez"
date: "09-10-2024"
output:
  xaringan::moon_reader:
    seal: false
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current% / 11" 
      highlightStyle: github
      highlightLines: true
      ratio: '16:9'
      countIncrementalSlides: false
      beforeInit: "macros.js"
---
```{r xaringan-themer, include=FALSE, warning=FALSE}
#this themer package allows us to change the .css themes through this chunk. Easier than changing the .css directly.
library(xaringanthemer)
style_mono_accent(
  base_color = "#7A003C", #maroon
  link_color = "#5E6A71", #grey
  text_bold_color = "#FDBF57", #gold
  text_slide_number_color = "#FDBF57", #gold
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

@@

- $i$ is a set of parcel point origin locations; $j$ is a set of destination locations.
- $O_j$ is a number of opportunities at $j$.
- $c_{ij}$ is the travel cost between $i$ and $j$.
- $f(\cdot)$ is an impedance function of $c_{ij}$; within the cumulative opportunity approach, it is a binary function that takes the value of 1 if $c_{ij}$ is less than a selected value

@@

### "Self Organinzing Maps" typologies: 15-min caring neighbourhoods

@@

- Lower income households tend to reside in $\color{Forestgreen}{\text{very completely caring}}$ or $\color{GreenYellow}{\text{completely caring}}$ neighbourhoods; 
- Those with a `r emo::ji("up_arrow")` proportion of children tend to reside in $\color{orange}{\text{somewhat caring - not complete}}$ neighbourhoods;
- Higher income households in DAs with a `r emo::ji("down_arrow")` of single-parent households tend to reside in $\color{red}{\text{not caring - not complete}}$ neighbourhoods.

@@

Who has resides in $\color{Forestgreen}{\text{very completely caring}}$ or $\color{GreenYellow}{\text{completely caring}}$ neighbourhoods?  (maybe people who cannot afford cars???)
- DAs that have lower median household incomes and a smaller children population (lower than 24% of the population are children 14 and under).
- and DAs which have especially high LICO are often in $\color{Forestgreen}{\text{very completely caring}}$ neighbourhoods.

Who more often resides in $\color{BurntOrange}{\text{somewhat caring - not complete}}$ neighbourhoods?  
- Of bottom income DAs, DAs with a high proportion of children population (higher than 24% of the population are children 14 and under).

Who more often resides in $\color{BrickRed}{\text{not caring - not complete}}$ neighbourhoods?  
- DAs of higher income DAs and with lower proportion of single-parent households (13% or lower)

@@

### Conclusions

@@

#### Implications of findings and future considerations:

- **Gentrification concerns**: populations residing in $\color{Forestgreen}{\text{very completely caring}}$  neighbourhoods are often lower-income and those in $\color{red}{\text{not caring - not complete}}$ neighbourhoods are better-off.

- **Potential to improve caring for children and children's self-maintenance**: 
  - Why do lower-income DAs with a `r emo::ji("up_arrow")` proportion of children, and DAs with `r emo::ji("up_arrow")` proportion single-parent households have (medium) $\color{orange}{\text{somewhat caring - not complete}}$ access? 
  - What barriers are they facing?
]

@@

#### Recap:

- **15 minute-city** is bridged with **Mobility of Care**: 15-minute caring neighbourhoods.
- A data-driven pattern recognition method (i.e., self-organizing maps) is used to create **15-minute caring neighbourhood typologies** using **accessibility** and **diversity** indicators.
- Used Self-Organizing Maps again to **profile** these typologies based on relevant **census variables**.
- Discussed implications: concerns about gentrification and mobility of care children's trips

@@

---
class: center, middle, inverse

## Thank you for listening! `r emo::ji("smile")`

### Any questions? Remarks?

-

#### Anastasia Soukhov, **soukhoa@mcmaster.ca** 
#### Léa Ravensbergen, Lucía Mejía Dorantes, Antonio Páez
